maze-dataset

### 2025-04-08 docs/paper/paper.tex
more implementation details, page flow

This package utilizes a simple, efficient representation of mazes. Using
an adjacency list to represent mazes would lead to a poor lookup time
for whether any given connection exists, while using an adjacency matrix
would waste memory by failing to exploit the structure (e.g., only 4 of
the diagonals would be filled in). Instead, we describe mazes with the
following representation: for a \(d\)-dimensional lattice with \(r\)
rows and \(c\) columns, we initialize a boolean array
\(A = \{0, 1\}^{d \times r \times c}\), which we refer to in the code as
a
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.connection_list}{\texttt{connection\_list}}.
The value at \(A[0,i,j]\) determines whether a downward connection
exists from node \([i,j]\) to \([i+1, j]\). Likewise, the value at
\(A[1,i,j]\) determines whether a rightwards connection to \([i, j+1]\)
exists. Thus, we avoid duplication of data about the existence of
connections, at the cost of requiring additional care with indexing when
looking for a connection upwards or to the left. Note that this setup
allows for a periodic lattice.

To produce solutions to mazes, two points are selected uniformly at
random without replacement from the connected component of the maze, and
the \(A^*\) algorithm (\protect\hyperlink{ref-A_star}{Hart et al.,
1968}) is applied to find the shortest path between them. The endpoint
selection can be affected by
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}},
and complications caused by this are detailed in
\autoref{success-rate-estimation}.

@@

>>>>

This package utilizes a simple, efficient representation of mazes as
subgraphs of a finite lattice, which we call a
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze}{\texttt{LatticeMaze}}.
Using an adjacency matrix would be memory inefficient by failing to
exploit the highly sparse structure -- for example, for a 2-dimsnionsal
maze, only 4 of the diagonals would be filled in. On the other hand,
using an adjacency list could lead to a poor lookup time for whether any
given connection exists.

Instead, we describe mazes with the following representation: for a
\(d\)-dimensional lattice with \(r\) rows and \(c\) columns, we
initialize a boolean array \[
  A = \{0, 1\}^{d \times r \times c}
\] which we refer to in the code as a
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.connection_list}{\texttt{connection\_list}}.
In two dimensions, the value at \(A[0,i,j]\) determines whether a
\emph{downward} connection exists from node \([i,j]\) to \([i+1, j]\).
Likewise, the value at \(A[1,i,j]\) determines whether a
\emph{rightward} connection to \([i, j+1]\) exists. Thus, we avoid
duplication of data about the existence of connections, at the cost of
requiring additional care with indexing when looking for a connection
upwards or to the left. Note that this setup allows for a periodic
lattice. This format generalizes to lattices of higher dimensions, but
our package provides only limited support for generation and
visualization of higher dimensional mazes. Generation of mazes is
detailed in
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}.

To produce solutions to mazes, two points are selected uniformly at
random without replacement from the connected component of the maze, and
the \(A^*\) algorithm (\protect\hyperlink{ref-A_star}{Hart et al.,
1968}) is applied to find the shortest path between them. The endpoint
selection can be controlled via
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}},
and complications caused by this are detailed in
\autoref{success-rate-estimation}.

@@

\newpage
